---
title: "Corpus Analysis"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  min_revdeps: 5
  min_coverage: .65
  package_size_outlier: 5000
---

```{r setup, include=FALSE}
library(cranlogs)
library(DT)
library(fs)
library(lubridate)
library(purrr)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)

Sys.setenv(
  TYPER_BASE_DIR=system2(
    "git", 
    c("rev-parse", "--show-toplevel"), 
    stdout=TRUE
  )
)

source("inc/paths.R")
source("inc/setup.R")
source("inc/latextags.R")

options(repos=CRAN_LOCAL_MIRROR)

create_tags(path(TAGS_DIR, "corpus.tex"), prefix="", default=TRUE)
```

## Load data

```{r}
pkgs <- read_csv(
  CRAN_PACKAGES_DETAILS,
  col_types=cols(
    package = col_character(),
    installed = col_logical(),
    loadable = col_logical(),
    version = col_character(),
    title = col_character(),
    size = col_double(),
    package_native_code = col_double(),
    package_r_code = col_double(),
    package_code = col_double(),
    runnable_code = col_double(),
    runnable_files = col_double(),
    runnable_code_examples = col_double(),
    runnable_code_vignettes = col_double(),
    runnable_code_tests = col_double(),
    runnable_files_examples = col_double(),
    runnable_files_vignettes = col_double(),
    runnable_files_tests = col_double(),
    coverage_expr = col_double(),
    coverage_line = col_double(),
    runnable_success = col_double(),
    runnable_error = col_double(),
    runnable_time = col_double(),
    runnable_rate = col_double(),
    funs_private = col_double(),
    funs_public = col_double(),
    funs = col_double(),
    revdeps = col_double(),
    funs_with_assert = col_double(),
    asserts = col_double(),
    uses_roxygen = col_logical()
  )
)
```

## Overview

### Thresholds

```{r}
overview_table(
  r("threshold code coverage", percent(params$min_coverage)),
  r("threshold revdeps", params$min_revdeps),
  r("package size outier", params$package_size_outlier)
)
```


### Table

```{r summary}
summary_corpus <- function(df, prefix) {
  overview_table(
    r(paste(prefix, "cran"), df),
    r(paste(prefix, "installed"), filter(df, installed)),
    r(paste(prefix, "loadable"), filter(df, loadable)),
    
    r(paste(prefix, "with runnable code"), filter(df, runnable_code > 0)),
    r(paste(prefix, "runnable code"), sum(df$runnable_code, na.rm=T)),
    r(paste(prefix, "examples code"), sum(df$runnable_code_examples, na.rm=T)),
    r(paste(prefix, "tests code"), sum(df$runnable_code_tests, na.rm=T)),
    r(paste(prefix, "vignettes code"), sum(df$runnable_code_vignettes , na.rm=T)),
    
    r(paste(prefix, "with coverage"), filter(df, coverage_expr > 0)),
    r(paste(prefix, "mean expr coverage"), percent(mean(df$coverage_expr, na.rm=T))),
    r(paste(prefix, "median expr coverage"), percent(median(df$coverage_expr, na.rm=T))),
  
    r(paste(prefix, "with R code"), filter(df, package_r_code > 0)),
    r(paste(prefix, "R code"), sum(df$package_r_code, na.rm=T)),
    r(paste(prefix, "with native code"), filter(df, package_native_code > 0)),
    r(paste(prefix, "native code"), sum(df$package_native_code, na.rm=T)),
    
    r(paste(prefix, "w asserts"), filter(df, asserts > 0)),
    r(paste(prefix, "asserts"), sum(df$asserts, na.rm=T)),
    r(paste(prefix, "funs with asserts"), sum(df$funs_with_assert, na.rm=T)),
    
    r(paste(prefix, "functions"), sum(df$funs, na.rm=T)),
    r(paste(prefix, "public functions"), sum(df$funs_public, na.rm=T)),
    r(paste(prefix, "private functions"), sum(df$funs_private, na.rm=T)),
    
    r(paste(prefix, "mean revdes"), mean(df$revdeps, na.rm=T)),
    r(paste(prefix, "median revdes"), median(df$revdeps, na.rm=T)),
    
    r(paste(prefix, "runtime"), as.duration(sum(df$runnable_time, na.rm=T))),
    r(paste(prefix, "mean runtime"), as.duration(mean(df$runnable_time, na.rm=T))),
    r(paste(prefix, "median runtime"), as.duration(median(df$runnable_time, na.rm=T))),
    
    r(paste(prefix, "using roxygen"), sum(df$uses_roxygen, na.rm=T))
  )
}

summary_corpus(pkgs, "All")
```

### Plots

```{r considered packages}
pkgs_considered <- filter(pkgs, loadable, package_r_code > 0, coverage_expr > 0)
```

```{r plot revdeps histogram}
pkgs_considered %>%
  ggplot(aes(x=revdeps)) + 
    geom_histogram(bins=50) +
    scale_x_log10(labels=scales::comma) +
    labs(x="# reverse dependencies (log)")
```

```{r plot coverage histogram}
pkgs_considered %>%
  ggplot(aes(x=coverage_expr)) + 
    geom_histogram() +
    scale_x_continuous(labels=scales::percent) +
    labs(x="Expression coverage")
```

```{r plot all considered packages}
plot_all <- function(df) {
  coverage_mean <- mean(df$coverage_expr)
  coverage_median <- median(df$coverage_expr)
  revdeps_mean <- mean(df$revdeps)
  revdeps_median <- median(df$revdeps)
  
  df %>%
    mutate(
      outlier=package_code > 5e4,
      label=if_else(outlier, str_glue("{package}"), as.character(NA))
    ) %>%
    ggplot(
      aes(
        x=coverage_expr, 
        y=revdeps,
        label=label,
        size=package_code
      )
    ) +
    geom_point(alpha=.6) +
    geom_text(size=3, check_overlap = T, vjust=1.5, na.rm = TRUE) + 
    
    geom_hline(aes(yintercept=revdeps_mean), linetype="dashed", color="black", size=0.2) +
    geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="black", size=0.2) +
    geom_vline(aes(xintercept=params$min_coverage),linetype="solid", color="red", size=0.4) +
    geom_hline(aes(yintercept=params$min_revdeps),linetype="solid", color="red", size=0.4) +
    
    scale_x_continuous(labels = scales::percent) + 
    scale_y_log10(labels = scales::comma) +
    scale_size(range = c(.1, 5), labels = scales::comma) +
      theme(
      legend.position=c(0.2, 0.85),
      legend.box="horizontal",
      legend.box.background = element_rect(fill="white", size=0.1)
    ) +
    labs(
      x="Code coverage", 
      y="Number of reverse dependencies (log)", 
      size="Code size",
      linetype=""
    ) +
    guides(
      size=guide_legend(order=1),
      color=guide_legend(order=2),
      linetype=F
    )
}

plot_all(pkgs_considered)
```

## Project corpus

We only consider packages that have at least `r params$min_revdeps` reverse dependencies and `r params$min_coverage*100`% expression coverage.

```{r project corpus}
corpus <- 
  filter(
    pkgs, 
    loadable, 
    runnable_code > 0, 
    revdeps >= params$min_revdeps, 
    package_r_code > 0, 
    coverage_expr >= params$min_coverage
  )

write_lines(corpus$package, CORPUS_FILE)
write_csv(corpus, CORPUS_DETAILS)
```

### Reverse dependencies

```{r}
pkgs_revdeps_raw <- read_csv(
  CRAN_REVDEPS, 
  col_types=cols(
    package = col_character(),
    revdep = col_character()
  )
)
corpus_revdeps <- pkgs_revdeps_raw %>% semi_join(corpus, by="package")
write_csv(pkgs_revdeps_raw, CORPUS_REVDEPS)
write_lines(unique(corpus_revdeps$revdep), CORPUS_REVDEPS_UNIQUE)
```

### Sanity check

TODO

### Overview

#### Table

```{r corpus summary}
summary_corpus(corpus, "corpus")

write_csv(corpus, CORPUS_DETAILS)
write_lines(corpus$package, CORPUS_FILE)
```

#### Plot

```{r plot corpus}
plot_corpus <- function(df, code_outlier) {
  coverage_mean <- mean(df$coverage_expr)
  coverage_median <- median(df$coverage_expr)
  revdeps_mean <- mean(df$revdeps)
  revdeps_median <- median(df$revdeps)
  
  df %>%
    mutate(
      outlier=package_code > code_outlier,
      label=if_else(outlier, str_glue("{package}"), as.character(NA))
    ) %>%
    ggplot(
      aes(
        x=coverage_expr, 
        y=revdeps,
        label=label,
        size=package_code
      )
    ) +
    geom_point(alpha=.6) +
    geom_text(size=3, check_overlap = T, vjust=1.5, na.rm = TRUE) + 
    
    geom_hline(aes(yintercept=revdeps_mean), linetype="dashed", color="red", size=0.2) +
    geom_vline(aes(xintercept=coverage_mean), linetype="dashed", color="red", size=0.2) +

    scale_x_continuous(limits = c(params$min_coverage, 1), labels = scales::percent) + 
    scale_y_log10(labels = scales::comma) +
    scale_size(range = c(.1, 3), labels = scales::comma) +
      theme(
      legend.position=c(0.2, 0.85),
      legend.box="horizontal",
      legend.box.background = element_rect(fill="white", size=0.1)
    ) +
    labs(
      x="Code coverage", 
      y="Number of reverse dependencies (log)", 
      size="Code size",
      linetype=""
    ) +
    guides(
      size=guide_legend(order=1),
      color=guide_legend(order=2),
      linetype=F
    )
}

plot_corpus(corpus, params$package_size_outlier)
ggsave(CORPUS_PLOT)
```

### List of corpus packages

```{r list of packages}
datatable(
  corpus, 
  filter="top", 
  options=list(
    pageLength=50,
    autoWidth=TRUE
  )
)
```

## Appendix: Most downloaded CRAN packages

```{r top cran downloads}
df_top_cran <- cran_top_downloads(when="last-month", count=100)
```

The cranlogs allow us to get 100 most downloaded packages over the last month.
How many of these popular packages do we have in out corpus?

```{r}
missed_top_cran_packages <- setdiff(df_top_cran$package, corpus$package)
```

```{r}
filter(df_top_cran, package %in% missed_top_cran_packages) %>%
  select(-from, -to) %>%
  datatable()
```

The main reason is that we use R 3.5. Some key dependencies such as Hmisc are no longer available.