---
title: Evaluation of Types
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
params:
  type_analysis_datafile: ../data/type-analysis/merged/merged.csv
  graph_dirpath: ../../paper/plots/
  table_dirpath: ../../paper/tables/
  macro_dirpath: ../../paper/macros/
---

```{r setup, echo=FALSE, eval=TRUE}

suppressPackageStartupMessages(library(stringr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(kableExtra))

macro_filename <- file.path(params$macro_dirpath, "macros.tex")

source("latex-macros.R")

save_graph <- function(plot, filename, device = cairo_pdf, width=12, height=4, ...) {
    ggsave(filename = file.path(params$graph_dirpath, filename), plot = plot, device = device, width=width, height=height, ...)
    plot
}

save_table <- function(content, filename) {
    write(content, file = file.path(params$table_dirpath, filename))
}

simplify_types <- function(type) {
    if(type %in% c("any", "externalptr", "expression", "pairlist")) {
        "any"
    }
    else {
        type
    }
}

is_polymorphic <- function(kind) {
    kind %in% c("...", "any", "list")
}

infer_kind <- function(type) {
    if(type %in% c("integer", "double", "character", "logical", "complex", "raw")) {
        "scalar"
    }
    else if(type %in% c("integer[]", "double[]", "character[]", "logical[]", "complex[]", "raw[]")) {
        "vector"
    }
    else if(type %in% c("^integer[]", "^double[]", "^character[]", "^logical[]", "^complex[]", 
                        "^integer", "^double", "^character", "^logical", "^complex")) {
        "^vector"
    }
    else if(str_detect(type, "^list")) {
        "list"
    }
    else if(str_detect(type, "^struct")) {
        "struct"
    }
    else if(str_detect(type, "^class")) {
        "class"
    }
    else if(type == "any") {
        "any"
    }
    else if(type == "null") {
        "null"
    }
    else if(type == "...") {
        "..."
    }
    else if(type == "environment") {
        "environment"
    }
    else {
        "other"
    }
}

type_analysis_data <- 
    read_csv(params$type_analysis_datafile,
             col_names = TRUE,
             col_types = cols(col_character(),
                              col_character(),
                              col_integer(),
                              col_integer(),
                              col_integer(),
                              col_character(),
                              col_integer()))

type_analysis_data <-
    type_analysis_data %>%
    mutate(type = map_chr(type, simplify_types)) %>%
    mutate(kind = map_chr(type, infer_kind))

to_percentage <- function(proportions) {
    str_c(round(100 * proportions, 2), "%")
}
```


# Incompatible Functions

```{r incompatible-functions, echo=FALSE, eval=TRUE}

###type_compatibility <-
###    type_analysis_data %>%
###    group_by(package_name, function_name, outer_alternative) %>%
###    summarize(parameter_count = length(unique(parameter_position))) %>%
###    ungroup() %>%
###    group_by(package_name, function_name) %>%
###    mutate(compatible = length(unique(parameter_count)) == 1) %>%
###    ungroup()
    
type_compatibility <-
    type_analysis_data %>%
    group_by(package_name, function_name) %>%
    summarize(compatible = length(unique(outer_alternative)) == 1) %>%
    ungroup()

incompatible_functions <-
    type_compatibility %>%
    filter(!compatible) %>%
    select(package_name, function_name) %>%
    distinct()

datatable(incompatible_functions)
```

There are `r nrow(distinct(incompatible_functions, package_name, function_name))` incompatible functions. 
We remove them from the rest of the analysis.

```{r remove-incompatible-functions, echo=FALSE, eval=TRUE}
type_analysis_data <-
    type_analysis_data %>%
    dplyr::filter(!(paste0(package_name, c("::"), function_name) %in%
                    paste0(incompatible_functions$package_name,
                           "::",
                           incompatible_functions$function_name)))
```

```{r high-level, echo=FALSE, eval=TRUE}
total_package_count <- nrow(distinct(type_analysis_data, package_name))
total_function_count <- nrow(distinct(type_analysis_data, package_name, function_name))
```

**Packages**: `r total_package_count`

**Functions**: `r total_function_count`


# Corpus

## Function Frequency

```{r function-frequency, eval=TRUE, echo=FALSE}
package_function_count <-
    type_analysis_data %>%
    group_by(package_name) %>%
    summarize(function_count = length(unique(function_name))) %>%
    ungroup()

datatable(package_function_count)

function_frequency <-
package_function_count %>%
mutate(function_count = as.integer(5 * ceiling(function_count / 5))) %>%
group_by(function_count) %>%
summarize(package_count = n()) %>%
ungroup() %>%
mutate(package_perc = round(100 * package_count / sum(package_count), 2)) %>%
mutate(package_cumperc = round(100 * cumsum(package_count) / total_package_count, 2))

datatable(function_frequency)

function_frequency %>%
ggplot(aes(function_count, package_count)) +
geom_col() +
scale_y_continuous(
    "Packages", 
    sec.axis = sec_axis(~ 100 * . / total_package_count, name = "Packages %")
  )
```

## Parameter Frequency

```{r parameter-frequency, eval=TRUE, echo=FALSE}
function_parameter_count <-
    type_analysis_data %>%
    group_by(package_name, function_name) %>%
    ## subtract one for return position (-1)
    summarize(parameter_count = length(unique(parameter_position)) - 1) %>%
    ungroup()

datatable(function_parameter_count)

parameter_frequency <-
function_parameter_count %>%
#mutate(parameter_count = as.integer(5 * ceiling(parameter_count / 5))) %>%
group_by(parameter_count) %>%
summarize(function_count = n()) %>%
ungroup() %>%
mutate(function_perc = round(100 * function_count / total_function_count, 2)) %>%
mutate(function_cumperc = round(100 * cumsum(function_count) / total_function_count, 2))

datatable(parameter_frequency)

parameter_frequency %>%
ggplot(aes(parameter_count, function_count)) +
geom_col() +
scale_y_continuous(
    "Function", 
    sec.axis = sec_axis(~ 100 * . / total_function_count, name = "Function %")
  )
```

# Monomorphism

```{r monomorphism, type-data, eval=TRUE,echo=FALSE}
monomorphic_data <-
type_analysis_data %>%
group_by(package_name, function_name, parameter_position) %>%
summarize(monomorphic = n() == 1) %>%
ungroup()

monomorphic_function_data <-
monomorphic_data %>%
group_by(package_name, function_name) %>%
summarize(monomorphic = all(monomorphic)) %>%
ungroup()

datatable(monomorphic_function_data)


monomorphic_function_count <- sum(monomorphic_function_data$monomorphic)
polymorphic_function_count <- sum(!monomorphic_function_data$monomorphic)

monomorphic_package_data <-
monomorphic_function_data %>%
group_by(package_name) %>%
summarize(monomorphism = round(100 * as.double(sum(monomorphic)) / length(monomorphic)),
          monomorphic = all(monomorphic)) %>%
ungroup() %>%
arrange(desc(monomorphism))


datatable(monomorphic_package_data)

monomorphic_package_count <- sum(monomorphic_package_data$monomorphic)
polymorphic_package_count <- sum(!monomorphic_package_data$monomorphic)

monomorphism_package_data <-
monomorphic_package_data %>%
group_by(monomorphism) %>%
summarize(package_count = n()) %>%
ungroup() %>%
arrange(desc(monomorphism)) %>%
mutate(package_cumsum = round(100 * cumsum(package_count) / total_package_count, 2))

datatable(monomorphism_package_data)

monomorphism_package_data %>%
ggplot(aes(monomorphism, package_count)) +
geom_col() +
scale_y_continuous(
    "Package", 
    sec.axis = sec_axis(~ 100 * . / total_package_count, name = "Package %")
  )
```

There are `r monomorphic_function_count` (`r round(100 * monomorphic_function_count/total_function_count, 2)`%) monomorphic functions.
There are `r monomorphic_package_count` (`r round(100 * monomorphic_package_count/total_package_count, 2)`%) monomorphic packages.

There are `r polymorphic_function_count` (`r round(100 * polymorphic_function_count/total_function_count, 2)`%) polymorphic functions.
There are `r polymorphic_package_count` (`r round(100 * polymorphic_package_count/total_package_count, 2)`%) polymorphic packages.

# Type Distribution

```{r type-distribution, type-data, eval=TRUE, echo=FALSE}
type_data <-
type_analysis_data %>%
select(kind, type) %>%
group_by(kind, type) %>%
summarize(type_count = n()) %>%
mutate(kind_count = sum(type_count)) %>%
ungroup() %>%
arrange(desc(kind_count, type_count)) %>%
mutate(kind_perc = round(100 * kind_count / sum(kind_count), 2),
       type_perc = round(100 * type_count / sum(type_count), 2)) %>%
mutate(kind_cumperc = round(100 * cumsum(kind_count) / sum(kind_count), 2),
       type_cumperc = round(100 * cumsum(type_count) / sum(type_count), 2))


datatable(type_data)


kind_count <-
type_analysis_data %>%
select(kind) %>%
group_by(kind) %>%
summarize(kind_count = n()) %>%
ungroup() %>%
arrange(desc(kind_count)) %>%
mutate(kind_perc = round(100 * kind_count / sum(kind_count), 2)) %>%
mutate(kind_cumperc = round(100 * cumsum(kind_count) / sum(kind_count), 2))


datatable(kind_count)


type_count <-
type_analysis_data %>%
select(type) %>%
group_by(type) %>%
summarize(type_count = n()) %>%
ungroup() %>%
arrange(desc(type_count)) %>%
mutate(type_perc = round(100 * type_count / sum(type_count), 2)) %>%
mutate(type_cumperc = round(100 * cumsum(type_count) / sum(type_count), 2))

datatable(type_count)
```

## Scalar Types

```{r scalar-type, eval=TRUE, echo=FALSE}
scalar_data <-
type_analysis_data %>%
filter(kind == "scalar") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))

datatable(scalar_data)
```


## Scalar-na Types

```{r scalar-na-type, eval=TRUE, echo=FALSE}
scalar_na_data <-
type_analysis_data %>%
filter(kind == "scalar-na") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))

datatable(scalar_na_data)
```

## Vector Types

```{r vector-type, eval=TRUE, echo=FALSE}
vector_data <-
type_analysis_data %>%
filter(kind == "vector") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))

datatable(vector_data)
```

## Vector-na Types

```{r vector-na-type, eval=TRUE, echo=FALSE}
vector_na_data <-
type_analysis_data %>%
filter(kind == "vector-na") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))

datatable(vector_na_data)
```

## Class Types

```{r class-type, eval=TRUE, echo=FALSE}
class_data <-
type_analysis_data %>%
filter(kind == "class") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))

datatable(class_data)
```

## List Types

```{r list-type, eval=TRUE, echo=FALSE}
list_data <-
type_analysis_data %>%
filter(kind == "list") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))


datatable(list_data)
```

## Struct Types

```{r struct-type, eval=TRUE, echo=FALSE}
struct_data <-
type_analysis_data %>%
filter(kind == "struct") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))


datatable(struct_data)
```

## Other Types

```{r other-type, eval=TRUE, echo=FALSE}
other_data <-
type_analysis_data %>%
filter(kind == "other") %>%
group_by(type) %>%
summarize(count = n()) %>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 2),
       cumulative_percentage = round(100 * cumsum(count) / sum(count), 2))


datatable(other_data)
```

## Union Types

```{r union-type, eval=TRUE, echo=FALSE}
union_data <-
type_analysis_data %>%
group_by(package_name, function_name, parameter_position) %>%
mutate(alternative_count = length(paste0(outer_alternative, "-", inner_alternative))) %>%
ungroup()
# %>%
#filter(alternative_count != 1)


union_frequency <-
union_data %>%
distinct(package_name, function_name, parameter_position, alternative_count) %>%
group_by(alternative_count) %>%
summarize(parameter_count = n()) %>%
ungroup() %>%
arrange(desc(parameter_count)) %>%
mutate(percentage = round(parameter_count / sum(parameter_count), 4),
       cumulative_percentage = round(cumsum(parameter_count) / sum(parameter_count), 4))

datatable(union_frequency)

total_parameter_count <- sum(union_frequency$parameter_count)

print(total_parameter_count)

union_frequency_folded <-
    union_frequency %>%
    select(alternative_count, parameter_count, percentage) %>%
    filter(alternative_count > 3) %>%
    summarize(alternative_count = "> 3", parameter_count = sum(parameter_count)) %>%
    mutate(percentage = round(as.double(parameter_count) / total_parameter_count, 3)) %>%
    print()
    
union_frequency_folded <-
    union_frequency %>%
    select(alternative_count, parameter_count, percentage) %>%
    filter(alternative_count <= 3) %>%
    mutate(alternative_count = as.character(alternative_count)) %>%
    rbind(union_frequency_folded)
    
datatable(union_frequency_folded)
    
(union_frequency_folded %>%
ggplot(aes(alternative_count, percentage)) +
geom_col() +
scale_y_continuous(sec.axis = sec_axis(~ . * total_parameter_count, labels = count_labels), labels = relative_labels) +
labs(x="Types", y="Parameters") +
scale_x_discrete(limits=c("> 3", 3, 2, 1)) +
coord_flip() +
theme(legend.title = element_blank(), legend.position = "top")) %>%
save_graph("union_frequency.pdf")


generate_macro("PercUnitypedPositions", 
               union_frequency_folded %>% filter(alternative_count == "1") %>% pull(percentage),
               macro_filename)

generate_macro("PercManytypedPositions", 
               union_frequency_folded %>% filter(alternative_count == "> 3") %>% pull(percentage),
               macro_filename)


union_types <-
union_data %>%
group_by(package_name, function_name, parameter_position, alternative_count) %>%
summarize(kind = paste0(sort(kind), collapse=" | ")) %>%
ungroup() %>%
group_by(alternative_count, kind) %>%
summarize(count = n()) %>%
ungroup() %>%
group_by(alternative_count) %>%
arrange(desc(count)) %>%
ungroup() %>%
group_by(alternative_count) %>%
mutate(percentage = round(100 * count / sum(count), 1),
       cumsum_percentage = round(100 * cumsum(count) / sum(count), 1)) %>%
ungroup() %>%
group_by(alternative_count) %>%
slice(1:20) %>%
ungroup()

datatable(union_types)

unityped_positions <-
union_types %>% 
filter(alternative_count == 1) 


polymorphic_unityped_rows <-
unityped_positions %>%
pull(kind) %>% 
map_lgl(is_polymorphic)

polymorphic_unityped_rows <- (1:nrow(unityped_positions))[polymorphic_unityped_rows]

print(polymorphic_unityped_rows)

unityped_positions %>%
select("Types" = kind, "Parameter #" = count, "%" = percentage, "Cumulative %" = cumsum_percentage) %>%
kable("latex", booktabs = T) %>%
kable_styling(full_width = F, latex_options = c("scale_down")) %>%
row_spec(polymorphic_unityped_rows, color = "black", background = "lightgray") %>%
save_table("uni-typed-breakdown.tex")


single_union_types <-
union_data %>%
filter(alternative_count == 1) %>%
group_by(kind) %>%
summarize(count = n())%>%
ungroup() %>%
arrange(desc(count)) %>%
mutate(percentage = round(100 * count / sum(count), 1),
       cumsum_percentage = round(100 * cumsum(count) / sum(count), 1))


datatable(single_union_types)
```

## 

```{r, eval = TRUE, verbose = FALSE}
union_frequency <-
union_data %>%
distinct(package_name, function_name, parameter_position) %>%
mutate(polymorphic = ifelse(alternative_count > 1, true, is_polymorphic(alternative_count)) %>%
group_by(package_name, function_name, polymorphic_position_count = )
group_by(alternative_count) %>%
summarize(parameter_count = n()) %>%
ungroup() %>%
arrange(desc(parameter_count)) %>%
mutate(percentage = round(parameter_count / sum(parameter_count), 4),
       cumulative_percentage = round(cumsum(parameter_count) / sum(parameter_count), 4))
```
