---
title: "Overview"
output: html_notebook
---

Ok.
First, load the full signatures, and make sure we source the data processing file.
There's an error here, not sure what's causing it.
This is a work-in-progress.

```{r}
source("../code/data_processing.R") # get the functions we need
```

```{r, echo=FALSE}
los <- list.files("../data/full_sigs", full.names=TRUE)
names <- list.files("../data/full_sigs")
names(los) <- substr(names, 1, nchar(names)-4)
los <- lapply(los, readRDS)
los <- remove_empties(los)
# annoying strange error that I should fix pops up
```

These are the updated counts.
```{r}
# argument counts
total_arguments <- apply_count_to_lopkg_c(los, args_count_all)
total_mono_type <- apply_count_to_lopkg_c(los, args_count_all_mono_type)
total_mono_attr <- apply_count_to_lopkg_c(los, args_count_all_mono_attr_pattern)
total_mono_clas <- apply_count_to_lopkg_c(los, args_count_all_mono_class)
total_mono_all <- apply_count_to_lopkg_c(los, args_count_all_mono_all)
```

```{r}
# function counts
total_functions <- apply_count_to_lopkg_c(los, length)
total_mono_type_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_type)
total_mono_attr_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_attr_pattern)
total_mono_clas_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_class)
total_mono_all_functions <- apply_count_to_lopkg_c(los, fun_count_all_mono_all)
```

This is how you get the top sigs.
```{r}
poly_type_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_type_args))
poly_type_args <- poly_type_args[order(-poly_type_args$count),] # sort by count
poly_type_args$perc <- round(poly_type_args$count/sum(poly_type_args$count) * 100, 2)
```

Then, pick out the ones you want.
The `format_df_for_print` function is just for letting the notebook display the data frame properly.
The type, attr, and class rows are lists of lists.
```{r}
format_df_for_print(poly_type_args[1:20,])
```

Similarly for attributes, and classes.
```{r}
poly_attr_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_attr_args))
poly_attr_args <- poly_attr_args[order(-poly_attr_args$count),] # sort by count
poly_attr_args$perc <- round(poly_attr_args$count/sum(poly_attr_args$count) * 100, 2)
```

```{r}
poly_class_args <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los, extract_poly_class_args))
poly_class_args <- poly_class_args[order(-poly_class_args$count),] # sort by count
poly_class_args$perc <- round(poly_class_args$count/sum(poly_class_args$count) * 100, 2)
```

Printing is as before.
Note that ATM I also have attribute type information in here.
I'll get it to display just the names as well.
I figured this might be interesting, and I don't expect many functions to be polymorphic in this way.
```{r}
format_df_for_print(poly_attr_args[1:20,])
```

```{r}
format_df_for_print(poly_class_args[1:20,])
```

How many things are just scalar?
```{r}
lo_scalars <- apply_row_extractor_to_lopkg_c(los, get_rows_from_df_only_scalar)
total_scalar_arguments <- apply_count_to_lopkg_c(lo_scalars, args_count_all)
```

Let's see what happens if we change type systems!
```{r}
los_basic <- lapply(los, function(lof) {lapply(lof, function(df) {change_type_systems_df(df, type_map_T0_to_r)})})
```

Let's get some counts.
This is for a type system which is in direct correspondence with R's typeof.
```{r}
# argument counts
total_mono_type_basic <- apply_count_to_lopkg_c(los_basic, args_count_all_mono_type)
total_mono_all_basic <- apply_count_to_lopkg_c(los_basic, args_count_all_mono_all)
```

What if we roll NULL into every type?
```{r}
los_basic_no_null <- lapply(los_basic, function(lof) {lapply(lof, function(df) {fold_NULL_into_other_types_df(df)})})
```

What are the counts now?
```{r}
total_mono_type_basic_no_null <- apply_count_to_lopkg_c(los_basic_no_null, args_count_all_mono_type)
total_mono_all_basic_no_null <- apply_count_to_lopkg_c(los_basic_no_null, args_count_all_mono_all)
```

Great.
Now, what are the top instances of type-based polymorphism?
```{r}
poly_type_args_basic_no_null <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los_basic_no_null, extract_poly_type_args))
poly_type_args_basic_no_null <- poly_type_args_basic_no_null[order(-poly_type_args_basic_no_null$count),] # sort by count
poly_type_args_basic_no_null$perc <- round(poly_type_args_basic_no_null$count/sum(poly_type_args_basic_no_null$count) * 100, 2)
```

What's the top?
```{r}
format_df_for_print(poly_type_args_basic_no_null[1:20,])
```

So, clearly, "real number polymorphism" is the top choice.
What if we fold ints and doubles together into reals?

```{r}
los_basic_no_null_real <- lapply(los_basic_no_null, function(lof) {lapply(lof, function(df) {fold_together_int_double_df(df)})})
```

What are the counts now?
```{r}
total_mono_type_basic_no_null_real <- apply_count_to_lopkg_c(los_basic_no_null_real, args_count_all_mono_type)
total_mono_all_basic_no_null_real <- apply_count_to_lopkg_c(los_basic_no_null_real, args_count_all_mono_all)
```

Great.
Now, what are the top instances of type-based polymorphism?
```{r}
poly_type_args_basic_no_null_real <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los_basic_no_null_real, extract_poly_type_args))
poly_type_args_basic_no_null_real <- poly_type_args_basic_no_null_real[order(-poly_type_args_basic_no_null_real$count),] # sort by count
poly_type_args_basic_no_null_real$perc <- round(poly_type_args_basic_no_null_real$count/sum(poly_type_args_basic_no_null_real$count) * 100, 2)
```

What's the top?
```{r}
format_df_for_print(poly_type_args_basic_no_null_real[1:20,])
```

Ok.
What if we want some per package information?
Well, to do it on one package (let's try rlang):

```{r}
rlang_tally <- tally_up_lofun_c(los$rlang)
rlang_tally$tot_args
rlang_tally$tot_mono_args_type
format_df_for_print(rlang_tally$top_poly_type)
```

As you'll see, the numbers here are low because we didn't actually trace rlang itself, just that some other packages used rlang, giving us limited results.

Now, how about if we take all of R's base classes (i.e., those which occur naturally in R) and lump them all into a "primitive" class.
What sort of class-based polymorphism do we see, now?

```{r}
los_primitive_class <- lapply(los, function(lof) {lapply(lof, class_to_primitive_df)})
```

```{r}
total_mono_clas_prim <- apply_count_to_lopkg_c(los_primitive_class, args_count_all_mono_class)
total_mono_clas_functions_prim <- apply_count_to_lopkg_c(los_primitive_class, fun_count_all_mono_class)
```

Let's get the top poly class now.
```{r}
poly_class_args_prim <- reduce_extracted_lopkgs_to_one(apply_extract_to_lopkg(los_primitive_class, extract_poly_class_args))
poly_class_args_prim <- poly_class_args_prim[order(-poly_class_args_prim$count),] # sort by count
poly_class_args_prim$perc <- round(poly_class_args_prim$count/sum(poly_class_args_prim$count) * 100, 2)
```

```{r}
format_df_for_print(poly_class_args_prim[1:20,])
```

The low percentage numbers suggest that no one user-defined pattern contributes too much to the class-based polymorphism.
