---
title: "Tracing R Base: +"
output: html_notebook
---

Preamble.

```{r}
source("../code/data_processing.R")
```

In this notebook, you'll find our results from tracing part of the R base package.

Unfortunately, tracing base isn't going to be quite so straightforward.
The first and most relevant issue is that a number of the base package's functions are primitives (i.e. builtins) and are not amenable to the Genthat framework.
I needed to transform the + function into a closure in order to get this to shake out, and needed to make a number of light modifications to Genthat altogether.

For the record, this is what I needed to do to get + to trace:

```{r}
# unlock binding so we can write to it
# unlockBinding(as.symbol("+"), baseenv())
# This makes + into a closure, so we can deal with it
# assign("+", function(e1, e2) .Primitive("+")(e1, e2), envir=baseenv())
# Actually decorate the base env functions
# genthat::decorate_function(`+`, env=baseenv())
```

The above takes place in the genthat-tracing-site-file.

To load the data, run:

```{r}
# technically in L2
# base_df <- readRDS("../data/some_plus_examples.RDS")
# base_df <- readRDS("../data/base_with_plus_and_brack.RDS")
base_df <- readRDS("../data/base_with_brack_fixed.RDS")

# for L0
base_df_L0 <- transform_new_df_w_TS_map(base_df, type_map_T0_to_r, unparam = T)
base_df_L0 <- unlist_new_df(base_df_L0)

# just types
base_df_just_types <- base_df %>% select(arg1_t, arg2_t, ret_t)
```

```{r}
# pass things over to java
# dont forget to tar

# save L0
write_csv(base_df_L0, "../data/plus_80_L0.csv")
# save L1
write_csv(base_df, "../data/plus_80_L1.csv")

# do the java

# read subtype
subtype_df_L0 <- read_csv("../data/plus_80_L0_subtype.csv.gz")
subtype_df_L0 <- subtype_df_L0[!subtype_df_L0$pkg == "SUBTYPE",]
subtype_df_L1 <- read_csv("../data/plus_80_L1_subtype.csv.gz")
subtype_df_L1 <- subtype_df_L1[!subtype_df_L1$pkg == "SUBTYPE",]
subtype_df_L2 <- read_csv("../data/plus_80_L1_L2.csv.gz")
subtype_df_L2 <- subtype_df_L2[!subtype_df_L2$pkg == "SUBTYPE",]
```

```{r}
# just get +

subtype_df_L0_p <- filter(subtype_df_L0, fun == "+") %>% salient_only_one_fun
subtype_df_L1_p <- filter(subtype_df_L1, fun == "+") %>% salient_only_one_fun
subtype_df_L2_p <- filter(subtype_df_L2, fun == "+") %>% salient_only_one_fun(just_types=F)

# just get [

subtype_df_L0_b <- filter(subtype_df_L0, fun == "[") %>% salient_only_one_fun(just_types = F)
subtype_df_L1_b <- filter(subtype_df_L1, fun == "[") %>% salient_only_one_fun(just_types = F)
subtype_df_L2_b <- filter(subtype_df_L2, fun == "[") %>% salient_only_one_fun(just_types=F)

```

[ is somewhat interesting.
There were some issues with [.
The number of arguments that it takes is somewhat related to the dimensions of the thing being indexed.
For example, in data frames you can have a call to [ with 3 arguments: the data frame first, followed by the row and then column numbers (optionally).
For lists, it's always just 2 arguments.
Not sure how ... best to deal with this.
Right now, I just have it take 2 arguments: the thing being accessed, and an index variable.

There is perhaps room for a more sophisticateed analysis here, looking at the domensions of the first argument.
I don't think it'll be too challenging, but again, do we want to go that far?
I think that would be useful for some IDE tool or something: for instance, telling the programmer that they aren't using the thing with the right amount of dimensions.
Then again, you can totally do that and it's completely fine, you just might not get what you expect.

Cool:
src/main/names.c contains ~ function headers for builtin and special functions (minus types).
Notably, it includes the arity of the functions.


